---
title: "final report"
author: "Kejing Yan"
date: "2023-05-03"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(formatR)
knitr::opts_chunk$set(echo = TRUE)
library(car)
#library(caret)
library(dplyr)
library(ggplot2)
library(hrbrthemes)
library(GGally)
library(gridExtra)
library(glmnet)
library(nlme)
library(MASS)
library(goftest)
library(lmtest)
library(plotmo)
theme_update(plot.title = element_text(hjust = 0.5))
opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
```

```{r}
# Processing data:
rm(list = ls())
# Read in data from Final Project Code.Rmd
bmw <- read.csv('~/BMWpricing_trainingvalidation.csv', as.is = T)
# Convert all continuous variables to numeric:
bmw$mileage <- as.numeric(bmw$mileage)
bmw$engine_power <- as.numeric(bmw$engine_power)
bmw$price <- as.numeric(bmw$price)

# Filter out implausible observations: 
bmw <- bmw %>% filter(mileage >= 0, )
bmw <- bmw %>% filter(engine_power > 0)

# Calculate the interval between sold and registration dates:
bmw.regis <- strptime(bmw$registration_date, "%m/%d/%Y")
bmw.sold <- strptime(bmw$sold_at, "%m/%d/%Y")
# Create age variable (in days)
bmw$age <- difftime(bmw.sold, bmw.regis, units = 'days')
bmw$age <- as.numeric(bmw$age)

# Convert all the categorical variables to factors:
colnames(bmw)
categorical.colnames <- colnames(bmw)[c(2:3, 6:16)]
for (.col in categorical.colnames){
bmw[, .col] <- factor(bmw[, .col], ordered = F)
}
```
```{r}
# Split data  into training and validation sets: 
DataSetTraining = bmw %>% filter(bmw$training == 1) # 50% of the data
DataSetValidation = bmw %>% filter(bmw$validation == 1) # 50% of the data

attach(DataSetTraining)
```

Scatter plot Matrix
```{r}
data <- data.frame(price, mileage, engine_power, age)
ggpairs(data, 
        columnLabels = c("Price", "Mileage", "Engine Power", "Age"),
        upper = list(continuous = wrap("points", alpha = 0.3, size=0.15),
                     combo = wrap("box", alpha = 0.8, size = 0.15, outlier.size=0.05)),
        lower = list(continuous = wrap('cor', size = 3),
                     combo = wrap("facethist", bins = 40)),
        axis.text.y = element_text(size = 6))

data.2 <- data.frame(price, sqrt(mileage), engine_power, log(age), sqrt(mileage)*log(age))
ggpairs(data.2, 
        columnLabels = c("Price", "sqrt(Mileage)", "Engine Power", "log(Age)", "Interaction Term"),
        upper = list(continuous = wrap("points", alpha = 0.3, size=0.15),
                     combo = wrap("box", alpha = 0.8, size = 0.15, outlier.size=0.05)),
        lower = list(continuous = wrap('cor', size = 3),
                     combo = wrap("facethist", bins = 40)),
        axis.text.y = element_text(size = 6))
```

look at sub-categories of model key and car type
```{r}
table(car_type)
table(model_key)
```

Categorization
```{r}
new.car.type <- list(
  others = c("coupe","subcompact","van", "convertible"),
  hatchback = c("hatchback"),
  estate  = c("estate"),
  sedan = c("sedan"),
  subcompact = c("subcompact"),
  suv = c("suv")
)

for (i in 1:length(new.car.type)) levels(DataSetTraining$car_type)[levels(DataSetTraining$car_type)%in%new.car.type[[i]]] <- names(new.car.type)[i]
table(car_type)

detach(DataSetTraining)
attach(DataSetTraining)
table(car_type)
```
```{r}
# Series 1:
levels(DataSetTraining$model_key)[levels(DataSetTraining$model_key)%in%c("114", "116", "118", "120", "123", "125", "135")] <- "series_1"
# Series 2: 
levels(DataSetTraining$model_key)[levels(DataSetTraining$model_key)%in%c("216 Active Tourer", "216 Gran Tourer", "218", "218 Active Tourer", "218 Gran Tourer", "220")] <- "series_2"
# Series 3:
levels(DataSetTraining$model_key)[levels(DataSetTraining$model_key)%in%c("316", "318 Gran Turismo", "318", "320 Gran Turismo", "320", "325", "325 Gran Turismo", "328", "330", "330 Gran Turismo", "335", "335 Gran Turismo")] <- "series_3"
# Series 4:
levels(DataSetTraining$model_key)[levels(DataSetTraining$model_key)%in%c("418 Gran Coupé", "420", "420 Gran Coupé", "425", "430", "430 Gran Coupé", "435", "435 Gran Coupé")] <- "series_4"
# Series 5:
levels(DataSetTraining$model_key)[levels(DataSetTraining$model_key)%in%c("518", "520", "520 Gran Turismo", "523", "525", "528", "530", "530 Gran Turismo", "535", "535 Gran Turismo")] <- "series_5"
# Series 6:
levels(DataSetTraining$model_key)[levels(DataSetTraining$model_key)%in%c("640", "640 Gran Coupé", "650")] <- "series_6"
# Series 7:
levels(DataSetTraining$model_key)[levels(DataSetTraining$model_key)%in%c("730", "740", "750")] <- "series_7"
# Series i:
levels(DataSetTraining$model_key)[levels(DataSetTraining$model_key)%in%c("i3", "i8")] <- "series_i"
# Series M:
levels(DataSetTraining$model_key)[levels(DataSetTraining$model_key)%in%c("M135", "M235", "M3", "M4", "M550")] <- "series_M"
# Series X:
levels(DataSetTraining$model_key)[levels(DataSetTraining$model_key)%in%c("X1", "X3", "X4", "X5", "X5 M", "X5 M50", "X6", "X6 M")] <- "series_X"
# Series Z:
levels(DataSetTraining$model_key)[levels(DataSetTraining$model_key)%in%c("Z4")] <- "series_Z"

table(DataSetTraining$model_key)
detach(DataSetTraining)
attach(DataSetTraining)
```
box plots for car types and model keys
```{r}
ggplot(DataSetTraining, aes(x = car_type, y = price)) + geom_boxplot() + xlab("Car Types") + ylab("Price Sold in 2018")
```
```{r}
ggplot(DataSetTraining, aes(x = model_key, y = price)) + geom_boxplot() + xlab("Model Keys") + ylab("Price Sold in 2018")
```
Feature 1-8
```{r}
par(mfrow = c(2,4))
# Box plots: 
ggplot(DataSetTraining, aes(x = feature_1, y = price)) + geom_boxplot() + xlab("Feature 1") + ylab("Price Sold in 2018")
ggplot(DataSetTraining, aes(x = feature_2, y = price)) + geom_boxplot() + xlab("Feature 2") + ylab("Price Sold in 2018")
ggplot(DataSetTraining, aes(x = feature_3, y = price)) + geom_boxplot() + xlab("Feature 3") + ylab("Price Sold in 2018")
ggplot(DataSetTraining, aes(x = feature_4, y = price)) + geom_boxplot() + xlab("Feature 4") + ylab("Price Sold in 2018")
ggplot(DataSetTraining, aes(x = feature_5, y = price)) + geom_boxplot() + xlab("Feature 5") + ylab("Price Sold in 2018")
ggplot(DataSetTraining, aes(x = feature_6, y = price)) + geom_boxplot() + xlab("Feature 6") + ylab("Price Sold in 2018")
ggplot(DataSetTraining, aes(x = feature_7, y = price)) + geom_boxplot() + xlab("Feature 7") + ylab("Price Sold in 2018")
ggplot(DataSetTraining, aes(x = feature_8, y = price)) + geom_boxplot() + xlab("Feature 8") + ylab("Price Sold in 2018")
```
Hypothesis testing in the mean for feature 1-8
```{r}
t.test(price[which(feature_1 == 'TRUE')], price[which(feature_1 == 'FALSE')])
t.test(price[which(feature_2 == 'TRUE')], price[which(feature_2 == 'FALSE')])
t.test(price[which(feature_3 == 'TRUE')], price[which(feature_3 == 'FALSE')])
t.test(price[which(feature_4 == 'TRUE')], price[which(feature_4 == 'FALSE')])
t.test(price[which(feature_5 == 'TRUE')], price[which(feature_5 == 'FALSE')])
t.test(price[which(feature_6 == 'TRUE')], price[which(feature_6 == 'FALSE')])
t.test(price[which(feature_7 == 'TRUE')], price[which(feature_7 == 'FALSE')])
t.test(price[which(feature_8 == 'TRUE')], price[which(feature_8 == 'FALSE')])
```
Exclude feature 7 in the following analysis
```{r}
# Fitting our full model:
formula_wo_ft <- 'price ~ sqrt(mileage) + engine_power + log(age) + I(sqrt(mileage)*log(age)) + car_type + model_key + feature_1 + feature_2 + feature_3 + feature_4 + feature_5 + feature_6 + feature_8'

m.mls <- lm(formula_wo_ft, data = DataSetTraining)
summary(m.mls)
```

Diagnostic plots
```{r}
par(mfrow = c(2,2))
plot(m.mls)
```
remove point 38
```{r}
DataSetTraining <- filter(DataSetTraining, price!=95200)
detach(DataSetTraining)
attach(DataSetTraining)
```

```{r}
m.mls <- lm(price ~ sqrt(mileage) + engine_power + log(age) + I(sqrt(mileage)*log(age)) + car_type + model_key + feature_1 + feature_2 + feature_3 + feature_4 + feature_5 + feature_6 + feature_8, data = DataSetTraining)
par(mfrow = c(2,2))
plot(m.mls)
```
```{r}
# Replotting Standardized Residual Plot w/o outlier:
StanResMLS <- rstandard(m.mls)
dataMLS <- data.frame(DataSetTraining$price,StanResMLS)

ggplot() +
  geom_point(data=dataMLS, aes(x=DataSetTraining$price, y=StanResMLS, color = "MLS"), size = 0.5) +
  geom_hline(yintercept=4,color='blue') + geom_hline(yintercept=-4, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("steelblue")) +
  labs(y = "Standarized Residual", x = "Price") + ggtitle("Training Standarized Residuals Plot of Initial Model") + 
  scale_y_continuous(breaks = c(-4, 4, 10, 20))
```

```{r}
vif(m.mls)

m.mls1 <- lm(price ~ sqrt(mileage) + engine_power + log(age) + I(sqrt(mileage)*log(age)) + model_key + feature_1 + feature_2 + feature_3 + feature_4 + feature_5 + feature_6 + feature_8, data = DataSetTraining)
summary(m.mls1)
par(mfrow = c(2,2))
plot(m.mls1)
```

```{r}
# BoxCox:
boxcox(m.mls1,plotit=TRUE)
```

```{r}
# Applying Transformation:
m.trans <- lm(sqrt(price) ~ sqrt(mileage) + engine_power + log(age) + I(sqrt(mileage)*log(age)) + model_key + feature_1 + feature_2 + feature_3 + feature_4 + feature_5 + feature_6 + feature_8, data = DataSetTraining)
summary(m.trans)
```
```{r}
par(mfrow = c(2,2))
plot(m.trans)
```
```{r}
# Standardized Residual Plot:
StanResMLS <- rstandard(m.trans)
newprice <- sqrt(price)
dataMLS <- data.frame(newprice,StanResMLS)
ggplot() +
  geom_point(data=dataMLS, aes(x=newprice, y=StanResMLS, color = "MLS"), size = 0.5) +
  geom_hline(yintercept=4,color='blue') + geom_hline(yintercept=-4, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("steelblue")) +
  labs(y = "Standarized Residual", x = "Transformed Price") + ggtitle("Standarized Residuals MLS Plot") + 
  scale_y_continuous(breaks = c(-4, 4, 10, 20))
```

WLS
```{r}
#create weight vector
sdprice.by.log.age <- by(price, log(age), sd)
# Getting mean:"
mprice.by.log.age <- by(price, log(age), mean)
# Plotting SD Price vs. Mean log(age):
plot(mprice.by.log.age, sdprice.by.log.age, xlab="Log(age)", ylab="SD of Price",
     main="SD Price versus log(age) mean; LOESS fit",col="red", lty=0, cex=1, pch=19)
# LOESS:
smooth.sd <- loess.smooth(mprice.by.log.age, sdprice.by.log.age)
# Plot LOESS regressor:
lines(smooth.sd, lty=2)
# Let's use the LOESS fit to create weights for WLS regression:
# Create LOESS object:
smooth.pricesd.orig <- loess(sdprice.by.log.age ~  mprice.by.log.age, control=loess.control(surface="direct"))
# Predict SD
# LOESS interpolant:
sdprice.fits <- predict(smooth.pricesd.orig,data.frame(mprice.by.log.age = log(age)))
my.weights1 <- 1/(sdprice.fits)^2
```


```{r}
m.trans1 <- lm(sqrt(price) ~ sqrt(mileage) + engine_power + log(age) + I(sqrt(mileage)*log(age)) + model_key + feature_1 + feature_2 + feature_3 + feature_4 + feature_5 + feature_6 + feature_8, data = DataSetTraining, weights = my.weights1)
summary(m.trans1)
```
```{r}
par(mfrow = c(2,2))
plot(m.trans1)
```

```{r}
# Standardized Residual Plot: 
StanResMLS <- rstandard(m.trans1)
newprice <- sqrt(price)
dataMLS <- data.frame(newprice,StanResMLS)
ggplot() +
  geom_point(data=dataMLS, aes(x=newprice, y=StanResMLS, color = "MLS"), size = 0.5) +
  geom_hline(yintercept=4,color='blue') + geom_hline(yintercept=-4, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("steelblue")) +
  labs(y = "Standarized Residual", x = "Transformed Price") + ggtitle("Training Standarized Residuals Plot of WLS Model") +
  scale_y_continuous(breaks = c(-4, 4, 10, 20))

```
Normality Test
```{r}
shapiro.test(resid(m.trans1))
```
reject the null hypothesis of normality

Heteroscedasticity Test
```{r}
bptest(m.trans1)
```
null hypothesis that the data is homoscedastic


LASSO
```{r}
attach(DataSetTraining)
#standardize variables
age_s = scale(log(age))
engine_power_s = scale(engine_power)
mileage_s = scale(sqrt(mileage))

x.df <- model.matrix(lm(sqrt(price) ~ mileage_s + age_s + engine_power_s + model_key + I(age_s*mileage_s) + feature_1 + feature_2 + feature_3 + feature_4 + feature_5 + feature_6 + feature_8, weights = my.weights1))
mod <- glmnet(x = x.df, y = sqrt(price), alpha = 1)
#plot(mod, xvar = "lambda", label = TRUE)

plot_glmnet(mod, label=TRUE)

cv_lasso_fit <- cv.glmnet(x.df, y = sqrt(price), alpha = 1, nfolds = 5)
plot(cv_lasso_fit)

fit <- glmnet(x.df, sqrt(price), lambda = cv_lasso_fit$lambda.min, nfolds = 5, weights = my.weights1)
fit$beta

cv_lasso_fit$lambda.min
```
```{r}
avPlots(m.trans1)
```
Remove feature 3 and 5 according to avplots.

```{r}
# Fitting new model based on LASSO:
final <- lm(sqrt(price) ~ sqrt(mileage) + engine_power + log(age) + model_key + I(sqrt(mileage)*log(age)) + feature_1 + feature_2 + feature_4 + feature_6 + feature_8, weights = my.weights1, DataSetTraining)
summary(final)
```


```{r}
StanResMLS <- rstandard(final)
newprice <- sqrt(price)
dataMLS <- data.frame(newprice,StanResMLS)
ggplot() +
  geom_point(data=dataMLS, aes(x=newprice, y=StanResMLS, color = "MLS"), size = 0.5) +
  geom_hline(yintercept=4,color='blue') + geom_hline(yintercept=-4, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("steelblue")) +
  labs(y = "Standarized Residual", x = 'Transformed Price') + ggtitle("Training Standarized Residuals Plot of Final Model") +
  scale_y_continuous(breaks = c(-4, 4, 10, 20))
```

```{r}
# Standarized Residuals vs Fitted
Fitted = fitted(final)
dataMLSFitted <- data.frame(Fitted,StanResMLS)

# MLS Stan. Res. vs Fitted plot 
ggplot() +
  geom_point(data=dataMLSFitted, aes(x=Fitted, y=StanResMLS, color = "MLS"), size = 0.5) +
  geom_hline(yintercept=4,color='blue') + geom_hline(yintercept=-4, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("steelblue")) +
  labs(y = "Standarized Residual") + labs(x = "Fitted value") +
  ggtitle("Training Std. Residuals vs Fitted of Final Model") + 
  scale_y_continuous(breaks = c(-4, 4, 10, 20))
```

```{r}
# Diagnostic plots
par(mfrow=c(2,2))
plot(final)
```

Validation

```{r}
# Validation --------------------------------------------------------------------------------------
# Residuals for training data
detach(DataSetTraining)
ResMLS <- resid(final)
# Mean Square Error
mean((ResMLS)^2)
```


```{r}
# Series 1:
levels(DataSetValidation$model_key)[levels(DataSetValidation$model_key)%in%c("114", "116", "118", "120", "123", "125", "135")] <- "series_1"
# Series 2: 
levels(DataSetValidation$model_key)[levels(DataSetValidation$model_key)%in%c("216 Active Tourer", "216 Gran Tourer", "218", "218 Active Tourer", "218 Gran Tourer", "220")] <- "series_2"
# Series 3:
levels(DataSetValidation$model_key)[levels(DataSetValidation$model_key)%in%c("316", "318 Gran Turismo", "318", "320 Gran Turismo", "320", "325", "325 Gran Turismo", "328", "330", "330 Gran Turismo", "335", "335 Gran Turismo")] <- "series_3"
# Series 4:
levels(DataSetValidation$model_key)[levels(DataSetValidation$model_key)%in%c("418 Gran Coupé", "420", "420 Gran Coupé", "425", "430", "430 Gran Coupé", "435", "435 Gran Coupé")] <- "series_4"
# Series 5:
levels(DataSetValidation$model_key)[levels(DataSetValidation$model_key)%in%c("518", "520", "520 Gran Turismo", "523", "525", "528", "530", "530 Gran Turismo", "535", "535 Gran Turismo")] <- "series_5"
# Series 6:
levels(DataSetValidation$model_key)[levels(DataSetValidation$model_key)%in%c("640", "640 Gran Coupé", "650")] <- "series_6"
# Series 7:
levels(DataSetValidation$model_key)[levels(DataSetValidation$model_key)%in%c("730", "740", "750")] <- "series_7"
# Series i:
levels(DataSetValidation$model_key)[levels(DataSetValidation$model_key)%in%c("i3", "i8")] <- "series_i"
# Series M:
levels(DataSetValidation$model_key)[levels(DataSetValidation$model_key)%in%c("M135", "M235", "M3", "M4", "M550")] <- "series_M"
# Series X:
levels(DataSetValidation$model_key)[levels(DataSetValidation$model_key)%in%c("X1", "X3", "X4", "X5", "X5 M", "X5 M50", "X6", "X6 M")] <- "series_X"
# Series Z:
levels(DataSetValidation$model_key)[levels(DataSetValidation$model_key)%in%c("Z4")] <- "series_Z"

detach(DataSetTraining)
attach(DataSetTraining)
```

```{r}
# remove the data with new levels in model_key
diff_level <- DataSetValidation$model_key[! DataSetValidation$model_key %in% DataSetTraining$model_key]
DataSetValidation <- DataSetValidation %>% filter(! model_key %in% diff_level)
model_key <- as.factor(model_key)
output <- predict(final, se.fit = TRUE, newdata = DataSetValidation)
ResMLSValidation <- DataSetValidation$price - (output$fit)^2

#MSE
mean((ResMLSValidation)^2)
```

```{r}
# Relative Mean Square Error for validation data
mean((ResMLSValidation)^2) / mean((DataSetValidation$price)^2)

# The same, but subtracting the overall mean in the denominator:
mean((ResMLSValidation)^2) / mean((DataSetValidation$price-mean(DataSetValidation$price))^2)

# Create data frame with validation observation and prediction
test = data.frame(DataSetValidation$price, output$fit^2, 1:length(output$fit));
colnames(test)[1] = "price"
colnames(test)[2] = "Prediction"
colnames(test)[3] = "Index"
# Plot GroundCO vs Prediction for Validation Data Set
ggplot(data = test, aes(x = Prediction, y = price)) + 
  geom_point() +
  ggtitle("Validation Price vs Prediction")
```

```{r}
# Further comparisons
ggplot(data = test, aes(x = Index)) +
  geom_line(aes(y = price, color = "Price")) +
  geom_line(aes(y = Prediction, color="Prediction"), linetype="twodash") +
  scale_color_manual(name = element_blank(), labels = c("Price","Prediction"),
                     values = c("darkred", "steelblue")) + labs(y = "") +
  ggtitle("Validation")

```

```{r}
# Hard to see, let's zoom in
test2 = test[200:250,]
# Plot GroundCO vs Prediction for Validation Data Set
ggplot(data = test2, aes(x = Index)) +
  geom_line(aes(y = price, color = "Price")) +
  geom_line(aes(y = Prediction, color="Prediction"), linetype="twodash") +
  scale_color_manual(name = element_blank(), labels = c("Price","Prediction"),
                     values = c("darkred", "steelblue")) + labs(y = "") +
  ggtitle("Validation")
```










